---
title: "Preliminary analysis and comparison of elicited prior distributions across conditions"
author: ""
date: "8/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gamlss)
library(tidyverse)
library(tidybayes)
library(magrittr)

theme_set( theme_light() ) +
theme_update(
  panel.border = element_blank(),
  panel.grid.major = element_line(colour = "#666666", size = 0.1),
  panel.grid.minor = element_line(colour = "#666666", size = 0.1),    
)
```

In this document we perform some descriptive analysis of the priors chosen by participants and compare them across different conditions.

## Data
First, we load the data downlaoded from qualtrics (this file is not shared as it contains identifiable information) and remove all the identifiers. We then save this deidentified file.

```{r data-clean, eval = FALSE}
data.raw = read_csv( file = "data/prior-setting-study.csv", 
                 col_types = list())

data.raw %>%
  rename( 
    t3 = `Duration (in seconds)`,
    `Confidence in answer` = Q1_1,
    `Previous stats analysis` = Q2,
    `Stats software` = Q3,
    `Knowledge software` = Q4,
    `Knowledge stats` = Q5,
    `Knowledge Bayesian` = Q6,
    Strategy = Q7,
    `Affect of choice` = Q8,
    Education = Q9,
    condition = cond
  ) %>%
  select(t3, ResponseId, `Confidence in answer`:Education, alpha:t2) %>%
  extract( 3:nrow(.), ) %>%
  mutate(
    t1 = as.numeric(t1),
    t2 = as.numeric(t2),
    t3 = as.numeric(t3),
    time = t1 + t2 + t3
  ) %>%
  select (t1, t2, t3, time, everything()) %>%
  write.csv( file = "data/prior-setting-study-deidentified.csv", row.names = FALSE )
```

## Load the data
We load the data and add the codes from the open code generation process.

```{r data}
data = read_csv( file = "data/prior-setting-study-deidentified.csv", 
                 col_types = list())

data %<>% mutate(
  WIP_alpha = c(NA, NA, TRUE, TRUE, FALSE, TRUE, FALSE, NA, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE, NA, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, NA, NA, FALSE, TRUE, FALSE, NA, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, TRUE, NA, NA, NA, TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, NA, NA, TRUE, NA),
  
  WIP_beta = c(NA, NA, FALSE, FALSE, TRUE, NA, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE, NA, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE, NA, FALSE, TRUE, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, NA, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, NA, NA, NA, FALSE, TRUE, TRUE),
  
  IP_alpha = c(NA, NA, FALSE, FALSE, TRUE, FALSE, TRUE, NA, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, NA, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, NA, NA, FALSE, FALSE, TRUE, FALSE, FALSE, TRUE, NA, FALSE, FALSE, TRUE, FALSE, NA, FALSE, NA, FALSE, TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, NA, FALSE, FALSE, NA, FALSE, FALSE, NA),
  
  IP_beta = c(NA, NA, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, NA, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, NA, TRUE, FALSE, FALSE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, NA, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, NA, NA, NA, FALSE, FALSE, FALSE)
) %>%
  drop_na( condition )

data %>%
  write.csv( file = "data/prior-setting-study-deidentified-2.csv", row.names = FALSE )
```

## Analysis

Check the number of responses in each condition

```{r}
data %>%
  group_by(condition) %>%
  summarise(n = n())
```

Calculate the amount of time spent by the participants in the  survey:

```{r}
data %>%
  mutate( time = (as.numeric(t1) + as.numeric(t2)) ) %>%
  mutate( t3 = (as.numeric(t3)) ) %>%
  summarise( 
    median = median(time), 
    sd = sd(time),
    median_qualtrics = median(t3),
    sd_qualtrics = sd(t3)
  )
```


### Experise

Visualize the responses to liker items

```{r}
data  %>%
  gather( domain, knowledge_level, c(`Knowledge stats`, `Knowledge Bayesian`, `Knowledge software`)) %>%
  mutate(
    `knowledge_level` = factor(`knowledge_level`, levels = c("Not knowledgeable at all", "Slightly knowledgeable", "Moderately knowledgeable", "Very knowledgeable", "Extremely knowledgeable") )
  ) %>%
  ggplot() +
  geom_bar(aes( knowledge_level )) + 
  scale_x_discrete( drop=FALSE, labels = function(x) sub(" "," \n", x, fixed=TRUE) ) +
  facet_grid(domain ~ .)
```

### Comparing the prior distributions
We first create a prior data frame which we will use to create visualizations of the prior density on the parameter scale, response scale and prior predictive density.

```{r}
priors <- data %>%
  select( ResponseId, condition, alpha, beta, WIP_alpha:IP_beta ) %>%
  gather( parameter, prior, alpha, beta ) %>%
  separate( col = prior, into = c("dist", "mu", "sigma"), sep = "([\\(,\\)])", extra = "drop" ) %>%
  mutate( 
    dist = replace(dist, dist == "N", "normal"),
    dist = replace(dist, dist == "t", "student_t")
  ) %>%
  mutate( mu = as.numeric(mu), sigma = as.numeric(sigma) ) %>%
  drop_na( condition ) %>%
  mutate( index = seq(1:nrow(.)))

get_density <- function( dist, mu, sigma, grid) {
  if (dist == "normal") {
    dnorm( grid, mu, sigma )
  } else if(dist == "student_t") {
    gamlss.dist::dTF( grid, mu, sigma, nu = 3)
  } else {
    NA
  }
}

get_density_response <- function( dist, mu, sigma, grid) {
  if (dist == "normal") {
    dnorm( log(grid), mu, sigma ) * abs( 1/grid)
  } else if(dist == "student_t") {
    gamlss.dist::dTF(log(grid), mu, sigma, nu = 3) * abs( 1/grid)
  } else {
    NA
  }
}
```

1. Probability density of priors on the intercept

```{r, priors-intercept, fig.height = 4, fig.width = 12}
priors.dens.alpha <- priors %>% 
  filter( parameter == "alpha" ) %>%
  mutate( grid = list(seq(-1, 7, by = 0.01)) ) %>%
  mutate( density = pmap(list(dist, mu, sigma, grid), ~ get_density(..1, ..2, ..3, ..4)) )

priors.dens.alpha %>%
  unnest( grid, density, .drop = FALSE) %>%
  ggplot() +
  geom_vline( xintercept = log(35), alpha = 0.5 ) +
  geom_line(aes( x = grid, y = density, color = dist, group = ResponseId), alpha = 0.7) +
  facet_wrap( ~ condition)
```

2. Probability density of priors on the intercept transformed to the response scale

```{r, priors-intercept-response, fig.height = 4, fig.width = 12}
priors.dens_response.alpha <- priors %>% 
  filter( parameter == "alpha" ) %>% 
  mutate( grid = list(seq(0.5, 140, by = 0.5)) ) %>%
  mutate( density = pmap(list(dist, mu, sigma, grid), ~ get_density_response(..1, ..2, ..3, ..4)) ) %>%
  unnest( grid, density, .drop = FALSE)
  
df.interval = tibble( x = c(24.6, 44.1), y = range(priors.dens_response.alpha$density)) %>%
  gather("axis", "value", x, y) %>%
  mutate( axis = c("x.lower", "x.upper", "y.lower", "y.upper")) %>%
  spread( axis, value)

p1 <- priors.dens_response.alpha %>%
  unnest( grid, density, .drop = FALSE) %>%
  ggplot() +
  geom_vline( xintercept = 128 ) +
  # geom_rect( data = df.interval, aes(xmin = x.lower, xmax = x.upper, ymin = y.lower, ymax = y.upper), fill = "grey50", alpha = 0.3 ) +
  geom_line(aes( x = grid, y = density, color = dist, group = ResponseId), alpha = 0.7) +
  scale_x_continuous( breaks = seq(0, 128, by = 16) ) +
  facet_grid( . ~ condition)
```

3. Histograms of the value of the location parameter of the priors set by participants in the different conditions

```{r, priors-intercept-mean-hist, fig.height = 4, fig.width = 12}
priors %>% 
  filter( parameter == "alpha" ) %>%
  ggplot() +
  geom_vline( xintercept = log(35), alpha = 0.5 ) +
  geom_histogram(aes( x = mu), color = "#ffffff", binwidth = 0.1) +
  scale_y_continuous( breaks = seq(0, 8, by = 2) ) +
  facet_wrap( ~ condition)
```

4. Histograms of the value of the scale parameter of the priors set by participants in the different conditions

```{r, priors-intercept-sd-hist, fig.height = 4, fig.width = 12}
priors %>% 
  filter( parameter == "alpha" ) %>%
  ggplot() +
  geom_histogram(aes( x = sigma), color = "#ffffff", binwidth = 0.1) +
  facet_wrap( ~ condition)
```

5. Probability density of the priors on the mean difference parameter, on the parameter scale

```{r, priors-meandiff, fig.height = 4, fig.width = 12}
priors.dens.beta <- priors %>% 
  filter( parameter == "beta" ) %>%
  mutate( grid = list(seq(-2, 2, by = 0.01)) ) %>%
  mutate( density = pmap(list(dist, mu, sigma, grid), ~ get_density(..1, ..2, ..3, ..4)) )

priors.dens.beta %>%
  unnest( grid, density, .drop = FALSE) %>%
  ggplot() +
  geom_vline( xintercept = 0, alpha = 0.5 ) +
  geom_line(aes( x = grid, y = density, color = dist, group = ResponseId), alpha = 0.7, size = 1.25) +
  facet_wrap( ~ condition)
```


5. Probability density of the priors on the mean difference parameter, on the multiplicative scale

```{r, priors-meandiff-response, fig.height = 4, fig.width = 12}
priors.dens.beta <- priors %>% 
  filter( parameter == "beta") %>%
  mutate( grid = list(seq(-2.5, 2.5, by = 0.01)) ) %>%
  mutate( density = pmap(list(dist, mu, sigma, grid), ~ get_density(..1, ..2, ..3, ..4)) )

priors.dens.beta %>%
  unnest( grid, density, .drop = FALSE) %>%
  ggplot() +
  geom_vline( xintercept = 0, alpha = 0.5 ) +
  geom_line(aes( x = grid, y = density, color = dist, group = ResponseId), alpha = 0.3, size = 1.25) +
  scale_x_continuous( breaks = log(c(0.1, 0.2, 0.4, 0.7, 1, 1.4, 2, 3, 5, 8)), labels = function(x) round(exp(x), 2) ) +
  facet_wrap( ~ condition)
```

6. Histogram of the values of the location parameter set for priors on the mean difference parameters

```{r, priors-meandiff-mean-hist, fig.height = 4, fig.width = 12}
priors %>% 
  filter( parameter == "beta" ) %>%
  ggplot() +
  geom_vline( xintercept = 0, alpha = 0.5 ) +
  geom_histogram(aes( x = mu, group = ResponseId), alpha = 0.7, binwidth = 0.1) +
  facet_wrap( ~ condition)
```

7. Histogram of the values of the scale parameter set for the priors on the mean difference parameter

```{r, priors-meandiff-sd-hist, fig.height = 4, fig.width = 12}
priors %>% 
  filter( parameter == "beta" ) %>%
  ggplot() +
  geom_histogram(aes( x = sigma, group = ResponseId), alpha = 0.7, binwidth = 0.1) +
  facet_wrap( ~ condition)
```


